---
title: "Analysis of Women income and education of Prestige"
author: "Jitong Jiang"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: default
params:
  state: "nc"
  download_raw_data: false
hitheme: tomorrow
highlighter: highlight.js
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 6,
  fig.path = "../results/"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Download data

The Prestige dataset, from the carData package, contains information on
102 occupations and 6 variables derived from Canadian census data and
social surveys conducted in the 1960s and early 1970s. The variables
include:

-   education: Average years of education of occupational incumbents in
    1971

-   income: Average income (in dollars) of incumbents in 1971

-   women: Percentage of incumbents who are women

-   prestige: Pineo–Porter prestige score based on a mid-1960s social
    survey

-   census: Canadian census occupational code

-   type: Occupational category (blue collar, white collar, or
    professional)

This analysis focuses on professional occupations only. The research
question is: Among **professional** occupations, how are the percentage
of women in an occupation and the average years of education related to
average income?

```{r}
if(!file.exists((here::here("data", "raw.RData")))){
  source(here("codes", "01_data_download.R"))
}else{
  load(here::here("data", "raw.RData"))
}
  
```

## Clean data

The Prestige dataset was loaded and restricted to professional
occupations only. Occupation names were converted from row names to a
variable, and only the variables relevant to the analysis (education,
income, and percentage of women) were retained. Observations with
missing values were removed to ensure complete data for all included
occupations.

```{r}
if(!file.exists((here::here("data", "clean.Rdata")))){
  source(here("codes", "02_data_cleaning.R")) 
}else{
  load(here::here("data", "clean.Rdata"))
}
Prestige
```

## Modeling

Let $Y_i$ denote the average income for occupation $i$,
$\text{Education}_i$ the average years of education, and
$\text{Women}_i$ the percentage of women in the occupation. The model is

$$
Y_i = \beta_0 + \beta_1 \,\text{Education}_i + \beta_2 \,\text{Women}_i + \varepsilon_i,
$$

where $\varepsilon_i$ are independent error terms with mean zero and
constant variance.

```{r}
if(!file.exists((here::here("data", "analysis.Rdata")))){
  source(here("codes", "03_data_analysis.R")) 
}else{
  load(here::here("data", "analysis.Rdata"))
}

summary(income_lm)
```

The linear regression model fitted with income as the response variable
and education and the percentage of women as predictors shows that the
women percantion in professional occupation is significantly negatively
associated with income, while education time shows a positive but
statistically insignificant association.

## Visualization

```{r}
source(here("codes", "04_data_visualization.R"))
```

```{r, income_vs_edu}
plot1
```

```{r, income_vs_women}
plot2
```

The scatter plots with fitted regression lines show a positive
association between education and income, and a negative association
between the percentage of women in and income. These are consist with
the results in the previous linear model.
